Doubly-Robust Inference in R using \\texttt{drtmle}

### 2022-06-25 paper/using_drtmle_obstudies.Rmd
update renv

An estimator of $\psi_0(a)$ may be motivated directly by (\ref{parameterDef}).
Suppose we have available an estimate of the OR, $\bar{Q}_n$, and an estimate of
the distribution function of baseline covariates, $Q_{W,n}$. An estimator of
$\psi_0(a)$ may be obtained by plugging these into (\ref{parameterDef}),
$\psi_n(a) = \int \bar{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical
distribution function of covariates is used so that this estimator can be
equivalently expressed
\begin{equation}
  \psi_{n,GCOMP}(a) = \frac{1}{n}\sum\limits_{i=1}^n \bar{Q}_n(a, W_i) \ .
  \label{eqn:gcomp}
\end{equation}
This estimator is commonly referred to as the G-computation (GCOMP) estimator.
Inference for GCOMP estimators based on parametric OR estimators may be
facilitated through the nonparametric bootstrap. However, if adaptive
approaches are used to estimate the OR, inference may not be available without
further modification to the estimator (e.g., undersmoothing as discussed above).

@@

An alternative estimator of the treatment-specific population-level mean may be
obtained by noting that (\ref{eqn:gcomp}) can be equivalently written as
\begin{align}
  E_0\{\bar{Q}(a,W)\} &= E_0\{E_0(Y \mid A = a, W)\} \notag \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A = a,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} Y \biggr\} \ .
          \label{iptwID}
\end{align}
We use $g_0(a \mid W) \coloneqq Pr_0(A = a \mid W)$ to denote the propensity
score. Suppose we had available $g_n$, an estimate of the PS. Equation
(\ref{eqn:iptwID}) motivates the estimator
\begin{equation*}
  \psi_{n,IPTW}(a) \coloneqq \frac{1}{n}\sum\limits_{i=1}^n
    \frac{I(A_i = a)}{g_n(a \mid W_i)} Y_i \ ,
\end{equation*}
which is referred to as the inverse probability of treatment weighted (IPTW)
estimator. Like GCOMP estimators, inference for IPTW estimators based on
parametric PS estimators may be facilitated through the nonparametric bootstrap;
however, unlike their GCOMP counterparts, IPTW estimators, being Z-estimators,
are constructed as solutions to an IPTW estimating equation, the empirical
variance of which may be used for standard error estimation. If adaptive
approaches are used to estimate the PS, inference may not be available without
further modification of the propensity score estimator, whether via targeting,
as we discuss below, or undersmoothing, as explored by
@ertefaie2022nonparametric and @hejazi2022efficient.

@@

This motivates a new class of estimators that combine both the GCOMP and IPTW
estimators. One example is the augmented IPTW (AIPTW) estimator
\begin{equation*}
\psi_{n,AIPTW}(a) = \psi_{n,IPTW}(a) + \frac{1}{n} \sum\limits_{i=1}^n \biggl\{
1 - \frac{I(A_i = a)}{g_n(a \mid W_i)} \biggr\} \  \bar{Q}_n(a, W_i) \ ,
\end{equation*}
which adds an augmentation term to the IPTW estimator. Equivalently, the
estimator can be viewed as adding an augmentation to the GCOMP estimator
\begin{equation} \label{oneStepEst}
\psi_{n,AIPTW}(a) = \psi_{n,GCOMP}(a) + \frac{1}{n} \sum\limits_{i=1}^n
\frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i - \bar{Q}_n(a,W_i)\} \ .
\end{equation}
The AIPTW estimator overcomes the two limitations of the GCOMP and IPTW
estimators. The estimator exhibits _double robustness_ in that it is consistent
for $\psi_0(a)$ if either $\bar{Q}_n$ is consistent for the true OR *or* $g_n$
is consistent for the true PS. Furthermore, if both the OR and PS are
consistently estimated and regularity conditions hold, one can establish a
normal limiting distribution for the AIPTW estimator. Simple estimators of the
variance of this limiting distribution can be derived that may be used to
construct Wald-style confidence intervals and hypothesis tests.

>>>>

An estimator of $\psi_0(a)$ may be motivated directly by (\ref{parameterDef}).
Suppose we have available an estimate of the OR, $\bar{Q}_n$, and an estimate of
the distribution function of baseline covariates, $Q_{W,n}$. An estimator of
$\psi_0(a)$ may be obtained by plugging these into (\ref{parameterDef}),
$\psi_n(a) = \int \bar{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical
distribution function of covariates is used so that this estimator can be
equivalently expressed
\begin{equation}
  \psi_{n,\text{GCOMP}}(a) = \frac{1}{n}\sum\limits_{i=1}^n \bar{Q}_n(a, W_i)
  \ . \label{eqn:gcomp}
\end{equation}
This estimator is commonly referred to as the G-computation (GCOMP) estimator.
Inference for GCOMP estimators based on parametric OR estimators may be
facilitated through the nonparametric bootstrap. However, if adaptive
approaches are used to estimate the OR, inference may not be available without
further modification to the estimator (e.g., undersmoothing as discussed above).

@@

An alternative estimator of the treatment-specific population-level mean may be
obtained by noting that (\ref{eqn:gcomp}) can be equivalently written as
\begin{align}
  E_0\{\bar{Q}(a,W)\} &= E_0\{E_0(Y \mid A = a, W)\} \notag \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A = a,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} Y \biggr\} \ .
  \label{iptwID}
\end{align}
We use $g_0(a \mid W) \coloneqq Pr_0(A = a \mid W)$ to denote the propensity
score. Suppose we had available $g_n$, an estimate of the PS. Equation
(\ref{iptwID}) motivates the estimator
\begin{equation*}
  \psi_{n,\text{IPTW}}(a) \coloneqq \frac{1}{n}\sum\limits_{i=1}^n
    \frac{I(A_i = a)}{g_n(a \mid W_i)} Y_i \ ,
\end{equation*}
which is referred to as the inverse probability of treatment weighted (IPTW)
estimator. Like GCOMP estimators, inference for IPTW estimators based on
parametric PS estimators may be facilitated through the nonparametric bootstrap;
however, unlike their GCOMP counterparts, IPTW estimators, being Z-estimators,
are constructed as solutions to an IPTW estimating equation, the empirical
variance of which may also be used for standard error estimation. If adaptive
approaches are used to estimate the PS, inference may not be available without
further modification of the propensity score estimator, whether via targeting,
as we discuss below, or undersmoothing, as explored by
@ertefaie2022nonparametric and @hejazi2022efficient.

@@

This motivates a new class of estimators that combine both the GCOMP and IPTW
estimators. One example is the augmented IPTW (AIPTW) estimator
\begin{equation*}
\psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{IPTW}}(a) + \frac{1}{n} \sum_{i=1}^n
  \biggl\{1 - \frac{I(A_i = a)}{g_n(a \mid W_i)} \biggr\} \ \bar{Q}_n(a, W_i)
  \ ,
\end{equation*}
which adds an augmentation term to the IPTW estimator. Equivalently, the
estimator can be viewed as adding an augmentation to the GCOMP estimator
\begin{equation} \label{oneStepEst}
\psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{GCOMP}}(a) + \frac{1}{n} \sum_{i=1}^n
\frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i - \bar{Q}_n(a,W_i)\} \ .
\end{equation}
The AIPTW estimator overcomes the two limitations of the GCOMP and IPTW
estimators. The estimator exhibits _double robustness_ in that it is consistent
for $\psi_0(a)$ if either $\bar{Q}_n$ is consistent for the true OR *or* $g_n$
is consistent for the true PS. Furthermore, if both the OR and PS are
consistently estimated and regularity conditions hold, one can establish a
normal limiting distribution for the AIPTW estimator. Simple estimators of the
variance of this limiting distribution can be derived that may be used to
construct Wald-style confidence intervals and hypothesis tests.
